---
output: github_document
bibliography: '`r system.file("REFERENCES.bib", package="HBEST")`'
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HBEST

<!-- badges: start -->
<!-- badges: end -->

`HBEST` is an `R` package that implements the nonparametric hierarchical Bayesian spectral analysis method known as HBEST. The details of this method can be found in @lee_hierarchical_2025. A detailed example using the function `HBEST()` can be found in the vignette.

The GitHub repository contains a Dockerfile to allow users to run RStudio in a local browser with `HBEST` already installed.

## Installation

You can install the development version of HBEST from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("llrebecca21/HBEST")
```

```{r example}
library(HBEST)
```


## Functions

This package contains user functions `HBEST()` and `HBEST_fast()`. Both functions implement the same method, but the `HBEST_fast()` function has slightly faster implementation. Both functions are made available to users since `HBEST()` is the code used in the publication @lee_hierarchical_2025. The vignette `HBEST_tutorial` contains a simple walk through of the `HBEST()` function. The data generation functions are also available to users and details of these functions can be found in the vignette.

## Docker Instructions

Assuming the user as already installed Docker on their local computer, one can follow these steps to run RStudio through a browser with HBEST and other dependent libraries already installed.

1. Download the **entire** zipped repository `HBEST` from [https://github.com/llrebecca21/HBEST](https://github.com/llrebecca21/HBEST), i.e., Code > Download Zip. Save the zipped `HBEST` folder to someplace you can navigate to easily with terminal commands. 

2. In your local terminal, navigate to where the repository folder is saved. i.e., `cd yourfilepath`

3. Run: `docker build -t hbest-rstudio .` to build the Docker container locally. This may take a couple of minutes.

4. Run the following code below and it is IMPORTANT to change the *changeme* to a password you will remember.

```{r, engine = 'bash', eval = FALSE}
docker run -d \
  --name hbest-rs \
  -p 8787:8787 \
  -e PASSWORD='changeme' \
  hbest-rstudio
```


5. Open Docker Desktop and find the container running `hbest-rs`, click the port link `8787:8787`. It is important that you are using a browser that supports http connections. If the port 8787 is already taken on your local computer, you will need to go into the Dockerfile and modify the last line `EXPOSE 8787` to another port number.

6. Login to RStudio with:

  `username: rstudio`
  
  `password: the password you created when running step 4 above.`
  
7. Check the package is installed by running: 
```{r, eval = FALSE}
library(HBEST)
# Call the help page for the HBEST function
?HBEST::HBEST
```




