---
title: "Data Generation Tutorial"
output: rmarkdown::html_vignette
bibliography: '`r system.file("REFERENCES.bib", package="HBEST")`'
vignette: >
  %\VignetteIndexEntry{Data_generation_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(HBEST)
```

This package also contains some exported data generation functions to help users replicate the data generation seen in the paper @lee_hierarchical_2025. The main data generation function is `gen_data()`. This is a wrapper function that easily allows users to choose the data generation they would like and specify any additional required inputs. `gen_data()` calls any of these functions:

  * `generate_AR()`
  * `generate_AR1_vary()`
  * `generate_AR2_mixture()`
  * `generate_MA4()`
  * `generate_MA4vary()`
  * `generate_MA4varyType2()`

Details of each of these functions can be found in their respective help page.











